\vspace{1.0cm} \LARGE \bf fluB

### 2013-05-27 fluB.tex
Changing "subtype" to "lineage", a few extra changes.

Influenza B viruses are increasingly being recognized as major contributors to morbidity attributed to seasonal influenza. 
Currently circulating influenza B isolates are known to belong to two antigenically distinct lineages (B/Victoria and B/Yamagata subtypes) on the basis of haemagglutinin inhibition (HI) assays. 
Frequent reassortment of the two subtypes has been noted in the past, but the effects of these reassortments have not been investigated in much detail.

@@

Influenza B viruses are considered to be comprised of two antigenically distinct lineages - Victoria and Yamagata (referred to as Vic and Yam, respectively) which are named after the strains (B/Victoria/02/87 and B/Yamagata/16/88, respectively) with antigenically distinct haemagglutinin (HA) surface glycoproteins. 
The two HA lineages still co-circulate today and are thought to have shared a most recent common ancestor in xxxx. 
All other influenza B segments also show a split in xxxx (comment - our estimate \textit{circa} 1984, use a figure to show TMRCAs and rates for each segment?).

@@

\subsection*{Sequences and subtyping}
We compiled a dataset of 452 complete influenza B genomes from GISAID (comment - need acknowledgment tables). 
Each strain was subtyped (assigned to either Vic or Yam subtype) in each segment by making maximum likelihood trees and looking whether it fell within the subtree containing B/Victoria/02/87 or B/Yamagata/16/88 sequences with the exception of the NS segment (B/Victoria/02/87 was a reassortant and possessed a Yam lineage NS), where B/Norway/1/84 was considered as being representative of Victoria lineage.
Each strain was thus assigned 8 subtypes depending on the combination of lineages it belongs to in all trees (for example all segments except for NS in strain B/Victoria/02/87 belong to Vic lineage and can thus be represented as V,V,V,V,V,V,V,Y). 
Seven of these where then used as discrete states in each of the trees (e.g. PB1 tree had PB2, PA, HA, NP, NA, MP and NS as traits and V or Y as trait values) in BEAST (comment - reference BEAST). (comment - give figure with proportions of subtypes for each segment)

\subsection*{Building trees}
For each segment it was attempted to make 3 replicates with 200 million MCMC states (comment - only third replicates ran for that long) each under the HKY model of nucleotide substitutions, 3 codon partitions, robust counting (comment - reference robust counting)) under the GMRF Bayesian skyride demographic model (comment - reference skyride).
In addition we used "relaxed" tips (comment - \# of relaxed tips) for when the dates of isolation for strains were incomplete (i.e. when only year of isolation was available).
We inferred the ancestral subtypes of lineages in each segment by using discrete traits (asymmetric CTMC matrix with BSSVS (comment - cite Philippe)). Because the posterior set of trees are labelled, we can infer intersubtype reassortments whenever a trait transition is observed (i.e. Y to V or V to Y). 
In addition, by reconstructing the subtypes of all other genomic segments jointly we can infer co-reassortment events when trait transitions occur on the same nodes in each tree.
We use a measure of persistence which is defined as the ratio of distances (in units of time) between a reassortant node and its most recent tip and compare it to the mean of distances between all nodes and their respective most recent tips. 
If this statistic is below 1 a reassortment event is presumed to be disadvantageous, if it is above 1 it is presumed it is advantageous.

\subsection*{Building a consensus set of reassortment events}
We built a consensus set of reassortment events by going through each tree in the posterior set of trees, identifying trait changes which are indicators of reassortment (with a cutoff of 1989 to avoid including the initial split between Vic and Yam as a reassortment event), taking the persistence value as the time between the date of reassortment (assumed to have occurred somewhere along the branch connecting nodes with opposite trait values) and the date of the most recent tip descended from that node. 
This was repeated for all nodes on the tree and the values averaged to give the average persistence time of nodes in the tree. 
We also noted which tips were descended from reassortant nodes.

@@

\section*{Results}

\begin{figure}[h]
	\centering		
	\includegraphics[width=0.95\textwidth]{figures/InfB_diversityOverTime}
	\caption{\textbf{Provisional diversity over time plot (tip to tip rather than branch to branch).}}
\end{figure}

>>>>

Influenza B viruses are increasingly being recognized as major contributors to morbidity attributed to seasonal influenza. 
Currently circulating influenza B isolates are known to belong to two antigenically distinct lineages (referred to as B/Victoria and B/Yamagata subtypes (comment - is subtype okay to use here?)) on the basis of haemagglutinin inhibition (HI) assays. 
Frequent reassortment between the segments of these two lineages has been noted in the past (comment - cite Lindstrom), but the effects of these reassortments have not been investigated in much detail.

@@

Influenza B viruses are considered to be comprised of two antigenically distinct lineages - Victoria and Yamagata (referred to as Vic and Yam, respectively) which are named after two strains: B/Victoria/02/87 and B/Yamagata/16/88, respectively, which possess antigenically distinct haemagglutinin (HA) surface glycoproteins. 
The two HA lineages still co-circulate today and are thought to have shared a most recent common ancestor in xxxx. 
All other influenza B segments also show a split in xxxx (comment - our estimate \textit{circa} 1984, use a figure to show TMRCAs and rates for each segment?).

@@

\subsection*{Sequences and subtyping}
We compiled a dataset of 452 complete influenza B genomes from GISAID (comment - need acknowledgment tables). 
Each strain was assigned to a lineage (either Vic or Yam subtype) in each segment by making maximum likelihood trees and looking whether it fell within the subtree containing B/Victoria/02/87 or B/Yamagata/16/88 sequences with the exception of the NS segment (B/Victoria/02/87 was a reassortant and possessed a Yam lineage NS), where B/Norway/1/84 was considered as being representative of Victoria lineage.
Each strain was thus assigned 8 lineages depending on the combination of lineages it belongs to in all trees (for example all segments except for NS in strain B/Victoria/02/87 belong to Vic lineage and can thus be represented as V,V,V,V,V,V,V,Y). 
Seven of these where then used as discrete states in each of the trees (e.g. PB1 tree had PB2, PA, HA, NP, NA, MP and NS as traits and V or Y as trait values) in BEAST (comment - reference BEAST). (comment - give figure with proportions of subtypes for each segment)

\subsection*{Building trees}
For each segment it was attempted to make 3 replicates with 200 million MCMC states (comment - only third replicates ran for that long) each under the HKY model of nucleotide substitutions, 3 codon partitions, robust counting (comment - reference robust counting)) under the GMRF Bayesian skyride demographic model (comment - reference skyride).
In addition we used "relaxed" tips (comment - \# of relaxed tips) for when the dates of isolation for strains were incomplete (i.e. when only year of isolation was available).
We inferred the ancestral state of lineages in each segment by using discrete traits (asymmetric CTMC matrix with BSSVS (comment - cite Philippe)). Because the posterior set of trees are labelled, we can infer interlineage reassortments whenever a trait transition is observed (i.e. Y to V or V to Y). 
In addition, by reconstructing the ancestral state of all other genomic segments jointly we can infer co-reassortment events when trait transitions occur on the same nodes in each tree.
We use a measure of persistence which is defined as the ratio of distances (in units of time) between a reassortant node and its most recent tip (the event is sampled uniformly over the branch displaying a trait value that is different from its parent node) and compare it to the mean of distances between all nodes and their respective most recent tips. 
If this statistic is below 1 a reassortment event is presumed to be disadvantageous, if it is above 1 it is presumed it is advantageous.

\subsection*{Building a consensus set of reassortment events}
We built a consensus set of reassortment events by going through each tree in the posterior set of trees, identifying trait changes which are indicators of reassortment (with a cutoff of 1989 to avoid including the initial split between Vic and Yam as a reassortment event), taking the persistence value as the time between the date of reassortment and the date of the most recent tip descended from that node. 
This was repeated for all nodes on the tree and the values averaged to give the average persistence time of nodes in the tree. 
We also noted which tips were descended from reassortant nodes to identify the same events in different segments.

@@

\section*{Results}
Reassortment events that involve a basic polymerase (PB1 or PB2) or HA, but never all 3, have 95\% HPDs that overlap 1 with the exception of a stray PB2-NA-NS events which is near 1 and has very low posterior probability. (comment - need a good way of representing all consensus events)

\begin{figure}[h]
	\centering		
	\includegraphics[width=0.95\textwidth]{figures/InfB_diversityOverTime}
	\caption{\textbf{Provisional diversity over time plot (tip to tip rather than branch to branch) (comment - slightly unusual that NS has quite wide HPD intervals at a few places, even though posterior smears don't show a bimodal distribution).}}
\end{figure}
